WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2014-05-28 ms.tex
Fine-tune notation for BAYESTAR likelihood

\marginpar{Actually, this does not give exactly the same Fisher matrix. Explain.}%
%
Another way is to plug the signal model from \Eref{eq:signal-model} \emph{evaluated at the ML parameter estimates} into the single-detector likelihood in \Eref{eq:gaussian-likelihood-spa}:
%
\begin{eqnarray*}
	\fl p\left(\hat{\boldsymbol\theta}_i \middle| \boldsymbol\theta \right)
	&:=&
	p\left(x_i(\omega) = h_i(\omega; \hat{\boldsymbol\theta}_i)
		\middle| \boldsymbol\theta \right)
	\\\fl
	&\propto& \exp \left[
		- \frac{1}{2} \int_0^\infty \frac{\left|
			\frac{\hat{\rho}_i}{\sigma_i(\hat{\boldsymbol\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})
			- \frac{\rho_i}{\sigma_i(\boldsymbol\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in})
		\right|^2}{S_i(\omega)} \, d\omega
	\right].
\end{eqnarray*}
%
If we assume that $\hat{\boldsymbol\theta}_\mathrm{in} = \boldsymbol\theta_\mathrm{in}$, then this reduces to
%
\begin{equation}
	p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
		\middle| \rho_i, \gamma_i, \tau_i \right) \propto
	\exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
		+ \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
		\right\}
	\right]
\end{equation}
%
with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a(t; \boldsymbol\theta_\mathrm{in})$ defined as
%
\begin{equation}
	a(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{\sigma_i(\boldsymbol\theta_\mathrm{in})} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
\end{equation}

To assemble the joint likelihood for the whole network, we just form the product from the individual detectors:
%
\begin{equation}
	\fl p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
		\middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \propto
	\exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
		+ \sum_i \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
		\right\}
	\right].
\end{equation}
%
To marginalize over the coalescence phase, we can insert an arbitrary nuisance phase $\eta$ by making the replacement $\tilde{\gamma}_i \rightarrow \tilde{\gamma}_i + \eta$. Then integrating over $\eta$ from $0$ to $2\pi$ and suppressing normalization factors, we get the \ac{BAYESTAR} likelihood
%
\begin{equation}
	\mathcal{L}_\mathrm{BAYESTAR} = \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
		\left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
		\right|
	\right],
\end{equation}
%
where $I_0$ is a modified Bessel function of the first kind.

>>>>

\marginpar{Actually, this does not give exactly the same Fisher matrix. Explain.}%
%
Another way is to plug the signal model from \Eref{eq:signal-model} \emph{evaluated at the ML parameter estimates} into the single-detector likelihood in \Eref{eq:gaussian-likelihood-spa}:
%
\begin{eqnarray*}
	\fl p\left(\hat{\boldsymbol\theta}_i \middle| \boldsymbol\theta \right)
	&:=&
	p\left(x_i(\omega) = h_i(\omega; \hat{\boldsymbol\theta}_i)
		\middle| \boldsymbol\theta \right)
	\\\fl
	&\propto& \exp \left[
		- \frac{1}{2} \int_0^\infty \frac{\left|
			\frac{\hat{\rho}_i}{\sigma_i(\hat{\boldsymbol\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})
			- \frac{\rho_i}{\sigma_i(\boldsymbol\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in})
		\right|^2}{S_i(\omega)} \, d\omega
	\right].
\end{eqnarray*}
%
If we assume that $\hat{\boldsymbol\theta}_\mathrm{in} = \boldsymbol\theta_\mathrm{in}$, then this reduces to
%
\begin{equation}
	p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
		\middle| \rho_i, \gamma_i, \tau_i \right) \propto
	\exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
		+ \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
		\right\}
	\right]
\end{equation}
%
with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a_i(t; \boldsymbol\theta_\mathrm{in})$ defined as
%
\begin{equation}
	a_i(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\boldsymbol\theta_\mathrm{in})}} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
\end{equation}

To assemble the joint likelihood for the whole network, we just form the product from the individual detectors:
%
\begin{equation}
	\fl p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
		\middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \propto
	\exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
		+ \sum_i \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
		\right\}
	\right].
\end{equation}
%
To marginalize over the coalescence phase, we can insert an arbitrary nuisance phase $\eta$ by making the replacement $\tilde{\gamma}_i \rightarrow \tilde{\gamma}_i + \eta$. Then integrating over $\eta$ from $0$ to $2\pi$ and suppressing normalization factors, we get the \ac{BAYESTAR} likelihood
%
\marginpar{Different symbol? Is $\mathcal{L}_\mathrm{BAYESTAR}$ too grandiose?}
%
\begin{equation}
	\mathcal{L}_\mathrm{BAYESTAR} = \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
		\left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
		\right|
	\right],
\end{equation}
%
where $I_0$ is a modified Bessel function of the first kind.
